A Case of Genetic Dilution in Improved Maize: Estimating Comparative Advantage of Adoption in Ethiopia

### 2022-07-21 document/slides/SPIA Final Meeting/finalpresentation.tex

\documentclass{beamer}
\usepackage[utf8]{inputenc}

%\usepackage{utopia} %font utopia imported

@@

\author[et al]{Michuda, Chiarella, Barriga-Cabanillas, Correa, }

% {Oscar Barriga-Cabanillas\inst{1} \\
% (with Travis Lybbert\inst{2})}

\institute[UC Davis] % (optional)
{
%  \inst{1}%
  Agricultural and Resource Economics\\
  UC Davis
 }

@@

\begin{itemize}
    \item Struggled to provide an adequate and reliable food supply
    \item Breakthroughs in maize germplasm ($\uparrow$ yields, $\uparrow$ drought tolerance)
\end{itemize}

@@

\begin{enumerate}
    \item Market integration
    \item Access to supplementary inputs
    \item Agro-ecological considerations
\end{enumerate}

@@

\begin{enumerate}
    \item Adoption where  net benefits are the highest
    \item How adaptation strategies augment comparative advantage in adoption
    \item Drivers of dis-adoption of improved maize seed
\end{enumerate}

@@

%---------------------------------------------------------
\begin{frame}
\frametitle{Empirical Strategy I}

Correlated Random Coefficient Model and comparative advantage (Suri 2011; Tjernstr\"{o}m et al. 2020; Michler et al. 2019; Barriga et al. 2018)

\begin{itemize}
    \item Farmers face unobservable heterogeneous returns to adoption (comparative advantage)
\end{itemize}

\begin{center}
$$
    y_{it}= \delta + \beta h_{it} + \theta_{i} + \phi\theta h_{it} + \tau_{i} + \epsilon_{it}
$$  
\end{center}

\begin{itemize}
    \item $y_{it}$: yields or profits for household i at time t
    \item $h$:  indicator for adoption
    \item $\tau$:  household’s absolute advantage
    \item $\theta$: household’s comparative advantage
    \item $\phi$: Gains from adopting, relative to a household’s comparative advantage. 
\end{itemize}

@@

%---------------------------------------------------------
\begin{frame}
\frametitle{Empirical Strategy II and Extensions}

Advantages:
\begin{enumerate}
    \item Identification does not rely on the existence of a valid instrumental variable (linear projection of observed adoption history)
    \item Disentangles a household’s absolute advantage from comparative advantage in adopting. 
\end{enumerate}

Extending the model:

\begin{itemize}
    \item Implementing a nonparametric panel identification in a novel GRC (Group Random Coefficients)
    \item Modeling and incorporating time-varying characteristics 
    \item Complement ESS data with precipitation (CHIRPS) and temperature (CPC) data
\end{itemize}  
\end{frame}

%---------------------------------------------------------
%---------------------------------------------------------
\begin{frame}
\frametitle{Group Random Coefficients Results}

@@

@@

%---------------------------------------------------------
\begin{frame}
\frametitle{Discussion}

\begin{itemize}
    \item Negative returns to adoption from self-reported improved varieties.
    \item But positive returns to more genetically pure, DTMZ, CG- sourced and newer varieties.
    \item All DNA fingerprints are 70\% pure or higher. Control group has high share of false positives: no proper control. 
    \item Negative results from GRC mask a large heterogeneity in returns to adoption.
\end{itemize}

@@

>>>>

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage{threeparttable}
%\usepackage{utopia} %font utopia imported

@@

\author[et al]{Aleksandr Michuda \\ Cristina Chiarella \\Oscar Barriga-Cabanillas \\Juan Correa }

% \institute[UC Davis] % (optional)
% {
% %  \inst{1}%
%   Agricultural and Resource Economics\\
%   UC Davis
%  }

@@

\begin{itemize}
    % \item Struggled to provide an adequate and reliable food supply
    \item Breakthroughs in maize germplasm ($\uparrow$ yields, $\uparrow$ drought tolerance)
\end{itemize}

@@

\begin{enumerate}[ {[}1{]} ]
    \item Market integration
    \item Access to supplementary inputs
    \item Agro-ecological considerations
\end{enumerate}

@@

\begin{enumerate}[ {[}1{]} ]
    \item Explore heterogeneous comparative advantages through GRC: find puzzling no significant results.
    \item Compare definitions of improved seeds: self-reported, second generation or recycled, purity of genetic material, year of release, types and source.
    \item Low adoption of newer and purest varieties, which do return in higher yields. Poor farmers paying a premium for genetically diluted improved varieties.
\end{enumerate}

@@

%---------------------------------------------------------
\begin{frame}
\frametitle{Data}

\begin{itemize}
    % \item Struggled to provide an adequate and reliable food supply
    \item We focus on maize producing households 
    \item Self-reported use of improved seed varieties
    \item Compare self-reported and cropcut yields
\end{itemize}

\begin{figure}
    \centering
    \includegraphics[scale=.4]{results/figures/map_hhids.pdf}
\end{figure}

% \textit{Add interesting image}
\end{frame}

%---------------------------------------------------------
\begin{frame}
\frametitle{Empirical Strategy I: Finding the returns from adoption}

\textbf{Understand the returns of adopting improved varieties: }
\begin{itemize}
    \item Some farmers have higher returns to adoption (comparative advantage)
    \item How comparative advantage plays a role on using improved seed varieties.
\end{itemize}

% (Suri 2011; Tjernstr\"{o}m et al. 2020; Michler et al. 2019; Barriga et al. 2018)

We use a Correlated Random Coefficient Model:
\vspace{-2em}
\begin{center}
$$
    y_{it}= \delta + \Delta h_{it} + \theta_{i} + \phi\theta h_{it} + \tau_{i} + \epsilon_{it}
$$  
\end{center}

\begin{itemize}
    \item $y_{it}$: yields or profits for household i at time t
    \item $h$:  indicator for adoption
    \item $\tau$:  household’s absolute advantage
    \item \textbf{$\theta$: household’s comparative advantage}
    \item \textbf{$\Delta$: The marginal effect of adoption}
    \item $\phi$: Gains from adopting, relative to a household’s comparative advantage.
\end{itemize}

@@

%---------------------------------------------------------
\begin{frame}
\frametitle{Empirical Strategy I: Finding the returns from adoption}

Advantages:
\begin{enumerate}
    \item Identification does not rely on a IV 
    % (linear projection of observed adoption history)
    \item Disentangles absolute advantage from comparative advantage in adoption. ($\theta_i = \theta_A - \theta_N$)
\end{enumerate}

\input{results/tables/traj}

% Extending the model:

% \begin{itemize}
%     \item Implementing a nonparametric panel identification in a novel GRC (Group Random Coefficients)
%     \item Modeling and incorporating time-varying characteristics 
%     \item Complement ESS data with precipitation (CHIRPS) and temperature (CPC) data
% \end{itemize}  
\end{frame}

%---------------------------------------------------------
%---------------------------------------------------------
\begin{frame}
\frametitle{Empirical Strategy II: Returns to Different Types of Hybrid}
% \frametitle{Group Random Coefficients Results}
\textbf{Why farmers do \textit{(not)} adopt?}

\begin{enumerate}
    \item There are negative returns from adoption
    \item Comparative advantage is close to 0 unless you always adopt (perhaps driven by 
\end{enumerate}

\begin{figure}
    \centering
    \label{fig:traj_sum}
    \includegraphics[scale=.4]{results/figures/traj_sum.png}
\end{figure}

@@

Why do we see farmers adopting despite the seemingly small negative returns?

\begin{itemize}
    \item We use the DNA-fingerprinting to shed light on this puzzle
    \item DNA allows us to assess what varieties (and purity) farmers are using 
    \item Unfortunately, only available in the last round
\end{itemize}

We implement a switching regression approach

@@

%---------------------------------------------------------
\begin{frame}
\frametitle{Discussion}
Negative and small returns to adoption from self-reported improved varieties.
\begin{itemize}
    \item But positive returns to more genetically pure, DTMZ, CG- sourced and newer varieties.
    \item All DNA fingerprints are 70\% pure or higher. Control group has high share of false positives: no proper control.

\end{itemize}
Negative GRC results mask a large heterogeneity in returns to adoption. 
\begin{itemize}
    \item Low purchase of newly bred hybrids, diluted DNA, low inputs use.

\end{itemize}

Effects of high-yielding varieties diminishes over time, but traditional varieties are well adapted to local agro-ecological conditions.

\begin{itemize}
    \item There may be a threshold in time above which it may be more profitable to use a traditional variety than an improved one that has been around for too many generations.

\end{itemize}

@@

%---------------------------------------------------------
\begin{frame}{}
    Thank you!
\end{frame}

\begin{frame}{Raw Trajectories}
    \begin{figure}
        \centering
        \includegraphics[scale=.5]{results/figures/theta.png}
    \end{figure}
\end{frame}

\bibliography{references}
